---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

# Reproducible Research: Peer Assessment 1
For background of project, please refer to [README.md](https://github.com/linearregression/RepData_PeerAssessment1/blob/master/README.md)

## Download data and unzip locally
```{r download, echo=TRUE}
zipData <- './activity.zip'
if(!file.exists("./activity.zip")) {
   print('Downloading activity.zip ....')   
   download.file(url='https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip', destfile=zipData, method='curl')
}
if(!file.exists("./activity.csv")) {
   print('Unzip activity.zip ....')  
   unzip(zipfile=zipData, overwrite=TRUE)
}
```
## Loading and preprocessing the data
Load data from csv
```{r loaddata, echo=TRUE}
Sys.setlocale(locale = "C")
fitdata<- read.csv(file='activity.csv', header=TRUE, stringsAsFactor=FALSE, na.strings = "NA", colClasses=c("integer", "Date", "integer"))

```
## What is mean total number of steps taken per day?
Group steps by date, then calculate mean number of steps.
Missing data is thrown away.
``` {r ggplot2lot, fig.width=4, fig.eight=3, message=FALSE, echo=TRUE}
library(plyr)
library(dplyr)
library(ggplot2)
fitdata <-na.omit(fitdata)
sumfitdata <- ddply(fitdata,~date, summarise, stepssum=sum(steps)) 
png('figure/stepsum.png', 480, 480)
p <- ggplot(data = sumfitdata, mapping = aes(x=date, y=stepssum)) +
     xlab('Date') +
     ylab('Total Steps') +
     ggtitle('Total Number Steps per Date')
p2 <- p +layer(geom = 'bar', geom_params = list(fill='green',color = 'steelblue'), stat = 'identity', stat_params = list(binwidth=0.5))
print(p2)
dev.off()
```
 - Mean nubmer of steps taken per day:
``` r
mean(sumfitdata$stepssum)
```   
 - Median number of steps taken per day:

``` r
median(sumfitdata$stepssum)
```

## What is the average daily activity pattern?

``` {r dailyactvity, echo=TRUE}
png('figure/stepdaily.png', 480, 480)
activitydata <- ddply(fitdata,c('interval'), summarise, 
     mean=mean(steps),
     sd = sd(steps)) 
p <- ggplot(data = activitydata, mapping = aes(x=interval, y=mean)) +
     xlab('Interval (min)') +
     ylab('Average Steps') +
     ggtitle('Daily Activity Pattern') 
p <- p+layer(geom="line", stat="identity")
print(p)
dev.off()

```
## Imputing missing values

Total number of missing data:

``` {r totalNA, echo=TRUE} 
sum(!complete.cases(fitdata))
```

## Are there differences in activity patterns between weekdays and weekends?

```r

```

